Automated topic naming to support cross-project analysis of software maintenance activities

### 2011-03-26 msr/msr2011.tex
10 pages

@@

@@

\vspace*{1em}
\section{Introduction}

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements. 
We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, http://dwheeler.com/sloccount.}. 
The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
MySQL contains $320,000$ lines of C and C++ source code.  
%Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily maintenance activities in our studies.

@@

%\noindent \textbf{Creating a validation corpus} --- 
\subsubsection{Creating a validation corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For MySQL 3.23
and MaxDB 7.500, the annotators (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations, but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly.
We looked at each period's topics, and assessed what the data ---
consisting of the frequency-weighted word lists and messages ---
suggested was the label for that topic. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label and sometimes they used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

%\todo[inline]{NEIL: abram, i got rid of the letter-grade thing because I'm worried people latch on to the idea we are C- students. I feel anything better than random is our criteria here}
ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
 reflecting how well a particular learner performed for the given data. 
ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). 
A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
The %area under the -- NE confusing to have area again ... and not importantana
ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.

We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).

@@

In this section we describe how to label topics based on dictionaries
mined from sources external to the projects themselves.

@@

Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519} .
That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
The labels we used:
\begin{quotation}
\small \noindent \textsf{
integrity, security,
interoperability, testability, maintainability, traceability,
accuracy, modifiability, understandability, availability, modularity,
usability, correctness, performance, verifiability, efficiency,
portability, flexibility, reliability.
}
\end{quotation}

@@

@@

Based on these results we find that \emph{reliability} and
\emph{usability} worked well for MaxDB in \textsf{exp2} and better in
\textsf{exp3}. 
\textsf{exp1} performed poorly.
MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
MySQL's results for \emph{efficiency} did not improve in \textsf{exp3}
but other qualities such as \emph{functionality} did improve. 
Many of ROC scores were $0.6$ or less, but our classifier nonetheless performed substantially better than random.

\subsection{Supervised labelling}
\label{sec:suplabelling}
Supervised labelling requires expert analysis of the correct
class/label to assign a label to a topic. In our approach, we use the top-level NFRs in the ISO9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.

We used a suite of supervised classifiers, WEKA~\cite{weka09}. 
WEKA includes machine learning tools such as support vector machines and Bayes-nets. 
We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}. %\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, which maps to what we observed while manually labelling topics.
The \emph{features} we used are word counts/occurrence  per topic, if
a word occurs frequently enough in a topic we consider it a feature of
the topic.

@@

\subsubsection{Analysis of the supervised labelling}
% \label{sec:suplabelling}
Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best performing learner per label. 
 We considered the best learner for a label to be the learner which
 had the highest ROC value for that label. 
The best learner is important because one uses a single learner per label. In applying our technique, we expect the user to select, for each NFR, the best learner possible.
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

@@

@@

@@

\section{Understanding software maintenance activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
%In order to test  our ROC values produced acceptable results, we now turn to
%\emph{applying} those techniques to understand software maintenance
%activities. 
%\todo[inline]{more on intro ... why should the reader care about
 % this. It needs to show how the stuff in the intro (applications) is
 % possible}
%%% XXX NEIL I don't get this.. VVVVV
%\todo[inline]{fix section headings}
As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
In this section we demonstrate the value of labelled topic extraction
in addressing this issue. 
% more why
Labelled topics address \emph{why}  because they show reasons why changes
occurred rather than \emph{how}. 
The \emph{how} of a change is the change itself, the purpose or
\emph{why} is what we are after.
We investigate the history of the two large-scale database systems that we studied. 
We use our technique to show the topic of development efforts over time in each project.
%We evaluated two research questions using the ground truth data that we annotated by hand:
We motivated our investigation with two research questions:
\begin{enumerate}
\item \emph{Do NFR frequencies change over time?} If a particular NFR was of more interest at one point in the life-cycle than another, this suggests that development activity shifted focus. For example, if a developer expected to see a recent focus on \emph{reliability}, but instead \emph{usability} dominated, they might re-prioritize up-coming work items.
\item \emph{Do projects differ in their relative interest in NFRs?} A
  project manager, especially a systems-manager, would be interested
  in knowing whether a particular NFR, such as \emph{reliability}, was
  more important for one project than another. This could be to
  confirm the initial design goals, or to track the progress on that
  quarter's objectives.  The difference in NFR proportion is
  interesting because it implies a difference in focus between two projects.
\end{enumerate}

@@

\noindent\textbf{1. Do NFR frequencies change over time?} --- 
%\subsubsection{Do label frequencies change over time?}
In both projects the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis timeframe for MaxDB being shorter than the timeframe for the MySQL product.

%\subsubsection{Do  projects differ in their relative  topic interest?}
\noindent\textbf{2. Do  projects differ in their relative  topic interest?} ---
Yes. MySQL 3.23 had proportionally more
topics labelled \emph{functionality}, while MaxDB had proportionally more
\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community, 
whereas MySQL was in its relative infancy, and  
security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers how often portability arises as a concern.
%On one hand, this analysis is unsurprising; on the other hand, it provides confirmation that our approach identifies actual patterns.
%\subsection{Summary}

@@

\subsection{Summary of techniques}
While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
supervised learners have the advantage of domain knowledge, which typically means improved results. 
% NEIL - we got criticized for making it sound like this is the obvious thing to do... and then why do you need a learner?
%Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period. 
Creating annotated topics (i.e., manual labels) for training is painstaking, but with a suitably representative set of topics, the effort is acceptable. To annotate \emph{all} topics took us approximately 20 hours per project, but we estimate only 10\% of the topics need annotation to produce useful results.

@@

\emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
We did not attempt to match our results to any particular model or
theory.
Also we did not have enough manpower to assess inter-rater reliability.

@@

>>>>

\usepackage{balance}

@@

% Only for camera ready -- prevent TeX from rendering widows
\widowpenalty=9999
\clubpenalty=9999

@@

\newpage

\section{Introduction}

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements. 
We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, \tiny{http://dwheeler.com/sloccount}.}. 
The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
MySQL contains $320,000$ lines of C and C++ source code.  
%Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily maintenance activities in our studies.

@@

%\noindent \textbf{Creating a validation corpus} ---

%FAT
\newpage

\subsubsection{Creating a validation corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For MySQL 3.23
and MaxDB 7.500, the annotators (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations, but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly.
We looked at each period's topics, and assessed what the data ---
consisting of the frequency-weighted word lists and messages ---
suggested was the label for that topic. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label and sometimes they used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

%\todo[inline]{NEIL: abram, i got rid of the letter-grade thing because I'm worried people latch on to the idea we are C- students. I feel anything better than random is our criteria here}
ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
 reflecting how well a particular learner performed for the given data. 
ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). 
A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
The %area under the -- NE confusing to have area again ... and not importantana
ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).

@@

In this section we describe how to label topics based on dictionaries
mined from sources external to the projects.

@@

Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519} .
That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
The labels we used:

%FAT we can remove this
\newpage
\begin{quotation}
\small \noindent \textsf{
integrity, security,
interoperability, testability, maintainability, traceability,
accuracy, modifiability, understandability, availability, modularity,
usability, correctness, performance, verifiability, efficiency,
portability, flexibility, reliability.
}
\end{quotation}

@@

%FAT
\balance
\newpage

@@

Thus we find that \emph{reliability} and
\emph{usability} worked well for MaxDB in \textsf{exp2} and better in
\textsf{exp3}. 
\textsf{exp1} performed poorly.
MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
MySQL's results for \emph{efficiency} did not improve in \textsf{exp3}
but other qualities such as \emph{functionality} did improve. 
Many of ROC scores were $0.6$ or less, but our classifier still performed substantially better than random.

%\vspace*{-1em}
\subsection{Supervised labelling}
\label{sec:suplabelling}
Supervised labelling requires expert analysis of the correct
class/label to assign a label to a topic. In our approach, we use the top-level NFRs in the ISO9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.

We used a suite of supervised classifiers, WEKA~\cite{weka09},
that includes machine learning tools such as support vector machines and Bayes-nets. 
We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}. %\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while manually labelling topics.
The \emph{features} we used are word counts/occurrence  per topic, if
a word occurs frequently enough in a topic we consider it a feature of
the topic.

@@

\subsubsection{Analysis of the supervised labelling}
% \label{sec:suplabelling}
Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best
performing learner per label: 
% We considered the best learner for a label to be 
the learner that had the highest ROC value for that label. 
The best learner is important because one uses a single learner per
label. When applying our technique, for each NFR one should select the best learner possible.
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

@@

\balance

@@

%FAT 
%\newpage

@@

\section{Understanding software maintenance activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
%In order to test  our ROC values produced acceptable results, we now turn to
%\emph{applying} those techniques to understand software maintenance
%activities. 
%\todo[inline]{more on intro ... why should the reader care about
 % this. It needs to show how the stuff in the intro (applications) is
 % possible}
%%% XXX NEIL I don't get this.. VVVVV
%\todo[inline]{fix section headings}
As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
In this section we demonstrate the value of labelled topic extraction
in addressing this issue. 
% more why
Labelled topics address \emph{why}  because they show reasons why changes
occurred rather than \emph{how}. 
The \emph{how} of a change is the change itself, the purpose or
\emph{why} is what we are after.
We investigate the history of the two large-scale database systems that we studied. 
We use our technique to show the topic of development efforts over time in each project.
%We evaluated two research questions using the ground truth data that we annotated by hand:
We motivated our investigation with two research questions:
\begin{enumerate}
\item \emph{Do NFR frequencies change over time?} If a particular NFR
  was of more interest at one point in the life-cycle than another,
  this suggests that development activity shifted focus. For example,
  if a developer expected to see a recent focus on \emph{reliability},
  but instead \emph{usability} dominated, they might re-prioritize
  upcoming work items.

\item \emph{Do projects differ in their relative interest in NFRs?} A
  project manager, especially a systems-manager, would be interested
  in knowing whether a particular NFR, such as \emph{reliability}, was
  more important for one project than another. This could be to
  confirm the initial design goals, or to track the progress on that
  quarter's objectives.  The difference in NFR proportion is
  interesting because it implies a difference in focus between two projects.
\end{enumerate}

%FAT

%\newpage

@@

\subsection{Do NFR frequencies change over time?}
%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
%\subsubsection{Do label frequencies change over time?}
In both projects the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

%FAT
\newpage

\subsection{Do  projects differ in their relative  topic interest?}
%\subsubsection{Do  projects differ in their relative  topic interest?}
%\noindent\textbf{2. Do  projects differ in their relative  topic interest?} ---
Yes. MySQL 3.23 had proportionally more
topics labelled \emph{functionality}, while MaxDB had proportionally more
\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community, 
whereas MySQL was in its relative infancy, and  
security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers how often portability arises as a concern.
%On one hand, this analysis is unsurprising; on the other hand, it provides confirmation that our approach identifies actual patterns.
%\subsection{Summary}

@@

%FAT
%\newpage
\subsection{Summary of techniques}
While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
supervised learners have the advantage of domain knowledge, which typically means improved results. 
% NEIL - we got criticized for making it sound like this is the obvious thing to do... and then why do you need a learner?
%Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period. 
Creating annotated topics (i.e., manual labels) for training is painstaking, but with a suitably representative set of topics, the effort is acceptable. To annotate \emph{all} topics took us approximately 20 hours per project, but we estimate only 10\% of the topics need annotation to produce useful results.

@@

\emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
We did not attempt to match our results to any particular model.
We unable to assess inter-rater reliability.

@@

\balance
\newpage
